---
title: "Storm Data Analysis Across United States: What Events are Accountable for Population Health and Econmic Losses? "
author: "Shuai"
date: "11/23/2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis
## Data Processing  
There are several steps we would like to follow:
1.  download and save the data as 'stormData.csv.bz2'.
```{r}
data_link<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(data_link,destfile = 'stormData.csv.bz2')
```
2.  The extracted csv file is as "stormData.csv". read the csv into data frame. Use data.table's fread() function to enable high efficiency csv reading.
```{r}
library(data.table)
library(R.utils)
if (!file.exists("stormData.csv")) {
bunzip2("stormData.csv.bz2", "stormData.csv", remove = FALSE, skip = TRUE)
} else {
  print("file is already downloaded!")
}
raw_data<-fread("stormData.csv",header=T)
```
3.  Perform exploratory data analysis. Use summary() function to explore parameters.
```{r}
summary(raw_data)
```
4.  Data Preparation.  
  To help analyze results directly related to required questions. I subset the original data into two indepndent dataset. The first question is : "Across the United States, which types of events (as indicated in the ð™´ðš…ðšƒðšˆð™¿ð™´ variable) are most harmful with respect to population health?". In order to answer what *"EVTYPE"* are most *"harmful"* related to *"population health"*. We need to find corresponding variables that are useful. Therefore, we subset the data (using subset() function to select variables of interest). The parameters of intertest are "EVTYPE, FATALITIES, INJURIES, REFUNM"
  4.1 Population Health Data Preparation
```{r}
PeopleHealth<-subset(raw_data,select=c(EVTYPE,FATALITIES,INJURIES,REFNUM))
PeopleHealthFull<-na.omit(PeopleHealth)#get full records
```
This PeopleHealthFull data is used for Results part.  
  4.2 Economic Loss Data Preparation
  Similar to 4.1, I subset the original data into a subset with variables of interest to answer questions -- "Across the United States, which types of events have the greatest economic consequences". These variables include:PROPDMG, PROPDMGEXP, CROPDMG
```{r}
econData<-subset(raw_data,select=c(EVTYPE,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP,REFNUM))
econDataFull<-na.omit(econData)
```
## Results  

### Across the United States, which types of events (as indicated in the ð™´ðš…ðšƒðšˆð™¿ð™´ variable) are most harmful with respect to population health?  
We use histogram to answer this question. Remember PeopleHealth data stores all information we need.
```{r}
PeopleHealth$People<-PeopleHealth$FATALITIES+PeopleHealth$INJURIES
summaryCount<-aggregate(People~EVTYPE,PeopleHealth,FUN=sum)
library(plyraltSummary<-aggregate(cbind(FATALITIES,INJURIES)~EVTYPE,PeopleHealth,FUN=sum)
)
summaryCount<-arrange(summaryCount,desc(People))
top10Health<-head(summaryCount,10)
```

Plot the top 10 event types that cause fatalites and injuries:
```{r}
library(ggplot2)
# Basic barplot
p<-ggplot(data=top10Health, aes(x=reorder(EVTYPE,-People), y=People))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab("Event Types")+
  ylab("Total Fatalities and Injuries")+
  ggtitle("Storms Events that are Most Harmful to Population Health")
p
```
### A+
  theme_light()cross the United States, which types of events have the greatest economic consequences?
```{r}
econDataFull$Loss<-as.numeric((econDataFull$PROPDMG))+as.numeric((econDataFull$CROPDMG))
summaryEcon<-aggregate(Loss~EVTYPE,econDataFull,FUN=sum)
library(plyr)
summaryEcon<-arrange(summaryEcon,desc(Loss))
top10Loss<-head(summaryEcon,10)
p<-ggplot(data=top10Loss, aes(x=reorder(EVTYPE,-Loss), y=Loss))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab("Event Types")+
  ylab("Total Crop Damages and Property Damages (kDollars)")+
  ggtitle("Storms Events Have the Greatest Economic Consequences")
p
```
## Co+
  theme_light()nclusion
Base on analysis between storm event types and 